Follow-up Samples

### 2016-12-22 submission/manuscript_outline_20161024.Rmd
update methods

***Study Design and Patient Sampling***
The sampling and design of the study was similar to that reported in Baxter, et al (**insert citation**).  In brief, study exclusion involved those who had already undergone surgery, radiation, or chemotherapy, had colorectal cancer before a baseline stool sample could be obtained, had IBD, a known hereditary non-polyposis colorectal cancer, or Familial adenomatous polyposis.  Samples used to build the model used for prediction were collected either prior to a colonoscopy or between 1 - 2 weeks after.  The bacterial microbiome has been shown to nomralize within this time period (**insert citation**).  Kept apart from this training set were a total of `r length(rownames(good_metaf))` individuals that not only had a sample as described previoulsy but also a follow up sample between `r min(good_metaf$time)` - `r max(good_metaf$time)` days after surgery and treatment had been completed. This study was approved by the University of Michigan Institutional Review Board.  All study participants provided informed consent and the study itself conformed to the guidelines set out by the Helsinki Declaration.

***Fecal Immunochemical Test and 16S rRNA Gene Sequencing***
FIT was analyzed as previously published using both OC FIT-CHEK and OC-Auto Micro 80 automated system (Polymedco Inc.) (**insert citation**). 16S rRNA gene sequencing was completed as previously described by Kozich, et al. (**insert citation**).  In brief, DNA extraction used the 96 well Soil DNA isolation kit (MO BIO Laboratories) and an epMotion 5075 automated pipetting system (Eppendorf).  The V4 variable region was amplified and the resulting product was split between three sequencing runs with control, adenoma, and carcinoma evenly represented on each run.  Each group was randomly assigned to avoid biases based on sample collection location.

@@

***Model Creation***
The Random Forest (**insert citation**) algorithm was used to create the model used for prediction of lesion (adenoma or carcinoma) for the `r length(rownames(good_metaf))` individuals with follow up samples.  The model included data on FIT, the bacterial microbiome, sex, age, Body Mass Index (BMI), whether the indivdiual was caucasion or not, history of cancer, and family history of cancer. Non-binary data was checked for near zero variance and auto correlation.  Data columns that had near zero variance were removed.  Columns that were correlated with each other over a Spearman correlation coefficient of 0.75 had one of the two columns removed.  This pre-processing was performed with the R package caret (v`r paste(packageVersion("caret"))`).  Optimiation of the mtry hyperparameter as well as data on the best, middle, and worse performance of the model involved taking the `r length(rownames(metaI))` samples and making 100 80/20 (train/test) splits in the data where control and lesion were equally represented in the 80 and 20 split, respectively.  This 80% portion was then split again into an 80/20 split, and run through 20 repeated 10-fold cross validations to optimzie the model's AUC (Area Under the Curve of the Receiver Operator Characteristic).  This resulting model was then tested on the 20% of the data that was originally held out from this overall process.  Once the ideal mtry was found the entire `r length(rownames(metaI))` sample set was used to create the final Random Forest model on which testing on the `r length(rownames(good_metaf))`-person cohort was completed.  The default cutoff of 0.5 was used as the threshold to classify individuals as positive or negative for lesion.

>>>>

***Study Design and Patient Sampling***
The sampling and design of the study was similar to that reported in Baxter, et al [@baxter_microbiota-based_2016].  In brief, study exclusion involved those who had already undergone surgery, radiation, or chemotherapy, had colorectal cancer before a baseline stool sample could be obtained, had IBD, a known hereditary non-polyposis colorectal cancer, or Familial adenomatous polyposis.  Samples used to build the model used for prediction were collected either prior to a colonoscopy or between 1 - 2 weeks after.  The bacterial microbiome has been shown to nomralize within this time period (**insert citation**).  Kept apart from this training set were a total of `r length(rownames(good_metaf))` individuals that not only had a sample as described previoulsy but also a follow up sample between `r min(good_metaf$time)` - `r max(good_metaf$time)` days after surgery and treatment had been completed. This study was approved by the University of Michigan Institutional Review Board.  All study participants provided informed consent and the study itself conformed to the guidelines set out by the Helsinki Declaration.

***Fecal Immunochemical Test and 16S rRNA Gene Sequencing***
FIT was analyzed as previously published using both OC FIT-CHEK and OC-Auto Micro 80 automated system (Polymedco Inc.) (**insert citation**). 16S rRNA gene sequencing was completed as previously described by Kozich, et al. [@kozich_development_2013].  In brief, DNA extraction used the 96 well Soil DNA isolation kit (MO BIO Laboratories) and an epMotion 5075 automated pipetting system (Eppendorf).  The V4 variable region was amplified and the resulting product was split between three sequencing runs with control, adenoma, and carcinoma evenly represented on each run.  Each group was randomly assigned to avoid biases based on sample collection location.

@@

***Model Creation***
The Random Forest [@breiman_random_2001] algorithm was used to create the model used for prediction of lesion (adenoma or carcinoma) for the `r length(rownames(good_metaf))` individuals with follow up samples.  The model included data on FIT and the bacterial microbiome. Non-binary data was checked for near zero variance and auto correlation.  Data columns that had near zero variance were removed.  Columns that were correlated with each other over a Spearman correlation coefficient of 0.75 had one of the two columns removed.  This pre-processing was performed with the R package caret (v`r paste(packageVersion("caret"))`).  Optimiation of the mtry hyperparameter as well as data on the best, middle, and worse performance of the model involved taking the `r length(rownames(metaI))` samples and making 100 80/20 (train/test) splits in the data where control and lesion were equally represented in the 80 and 20 split, respectively.  This 80% portion was then split again into an 80/20 split, and run through 20 repeated 10-fold cross validations to optimzie the model's AUC (Area Under the Curve of the Receiver Operator Characteristic).  This resulting model was then tested on the 20% of the data that was originally held out from this overall process.  Once the ideal mtry was found the entire `r length(rownames(metaI))` sample set was used to create the final Random Forest model on which testing on the `r length(rownames(good_metaf))`-person cohort was completed.  The default cutoff of 0.5 was used as the threshold to classify individuals as positive or negative for lesion.

***Analysis Overview***
